### 2012-08-02 Doc/VRIPHYS2012/vriphys2012.tex
proofread of section 4
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@449 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\usepackage{egweblnk}
\usepackage{cite}

@@

Extending the work of Comas \cite{Comas2010c} we have designed a shell element that uses B\'ezier interpolation polynomials.
Our element solves the problems of the element presented by Comas:
(i) it uses all 6 degrees of freedom,
(ii) the deformations are completely symmetric and
(iii) the boundary between elements is continuous.
We present a two stage interpolation:
(i) first a kinematic link between 6 DOF nodes of element and nodes of B \'ezier control mesh is defined and
(ii) interpolation function of cubic B\'ezier triangle is applied to interpolate on the surface of the element.
Mapping of high resolution mesh on the curved surface of the shell.
Finally from the kinematic link we derive formulas for
(i) interpolating velocity on the element surface and for
(ii) distribution of normal force applied on the element surface back to 6 DOF nodes of the element.
This allows us to use smooth mapped surface for collision detection and allows the propagation of proper collision response back onto the mechanical mesh.

@@

Because of the bending property of the shells relatively few elements are
necessary to simulate curved surface. To enrich the visual experience from
the simulated object it is desirable to use more triangles in the areas
with high curvature during rendering. Besides the visual accuracy the high
resolution mesh correctly modeling the curved areas is needed for
interaction and correct simulation of collisions and constraints. To handle
these we also need to know how to map the velocities from the mechanical
mesh onto the high resolution mesh and how to map forces acting on the high
resolution mesh back onto the mechanical mesh.

@@

For contact modeling it is often necessary to know the velocity of each point.
The values are directly available to us for the nodes of the mechanical mesh
but they can also be computed for any point on the surface.
Differentiating \eqref{eq-cubicbez2} by time we get:
%
\begin{equation}
    \dot{T} = \sum_{i=1}^{10} n_i \dot{P_i} = \sum_{i=1}^{10} n_i V_i
\end{equation}

Which means we use the same expression to interpolate the velocity only by
substituting the velocities at control points. For the corner nodes we
already have the velocities $V_i$. For the internal nodes we need to compute
them. We again make use of the kinematic link between nodes and solve the
problem of computing the velocity of a point attached to the moving rigid body.
For the edge points we have:
%
\begin{eqnarray}
    V_j = V_i + \omega_i \times P_{i-j}
\end{eqnarray}
%
where $i$ is the index of the corner node the edge node is attached to and
$\omega_i$ is the angular velocity at the node. For the central node
the formula is based on \eqref{eq-central2}:
%
\begin{eqnarray}
    V_i = \frac{1}{3} \sum_{i=1}^3 \left( V_i + \omega_i \times P_{i-10} \right)
\end{eqnarray}

\subsection{Projection of Normal Forces}

For physical interaction with the object one also requires that the normal
forces applied on the object's surface are transmitted back as forces
and torques on the vertices of the mechanical mesh.
We compute the influence of force $F$ on the corners through all the
associated control points by the formula:
%
\begin{eqnarray}
    F_1 & = & F \sum_{i \in \{1,4,5\}} n_i + \frac{n_{10}}{3} F \\
    F_2 & = & F \sum_{i \in \{2,6,7\}} n_i + \frac{n_{10}}{3} F \\
    F_3 & = & F \sum_{i \in \{3,8,9\}} n_i + \frac{n_{10}}{3} F 
\end{eqnarray}

Similarly we can compute the torques applied through edge control points and
the central control points:
%
\begin{gather}
    \begin{split}
        \theta_1 = & P_{1-4} \times (n_4 F) + P_{1-5} \times (n_5 F) + \\
        & + \frac{1}{3} P_{1-10} \times (n_{10} F)
    \end{split}
    \\
    \begin{split}
        \theta_2 = & P_{2-6} \times (n_6 F) + P_{2-7} \times (n_7 F) + \\
        & + \frac{1}{3} P_{2-10} \times (n_{10} F) \\
    \end{split}
    \\
    \begin{split}
        \theta_3 = & P_{3-8} \times (n_8 F) + P_{3-9} \times (n_9 F) + \\
        & + \frac{1}{3} P_{3-10} \times (n_{10} F) \\
    \end{split}
\end{gather}

@@

Because of the formulation of B\'ezier triangle is intrinsically present
there is a clean and obvious way of mapping high-resolution mesh for better
visualization. We also provide a technique of computing velocities of any point
on the surface and formula for back-projection of normal forces applied on the
surface onto the mechanical mesh. 
Those are essential components for the surgical simulators that need accurate contact geometry modeling without an overflow of computation need
The efficiency of the approach is demonstrated through convergence tests,  computation time benchmark and application to two different  simulation cases.

>>>>

\usepackage{egweblnk}
\usepackage{cite}
\usepackage{ulem}

@@

Extending the work of Comas \cite{Comas2010c} we have designed a shell element that uses B\'ezier interpolation polynomials.
Our element solves the problems of the element presented by Comas:
(i) it uses all 6 degrees of freedom,
(ii) the deformations are completely symmetric and
(iii) the boundary between elements is continuous.
We present a two stage interpolation:
(i) first a kinematic link between 6 DOF nodes of element and nodes of B \'ezier control mesh is defined and
(ii) interpolation function of cubic B\'ezier triangle is applied to interpolate on the surface of the element.
Mapping of high resolution mesh on the curved surface of the shell.
Finally from the kinematic link we derive formulas for
(i) interpolating velocity on the element surface and for
(ii) distribution of force applied on the element surface back to 6 DOF nodes of the element.
This allows us to use smooth mapped surface for collision detection and allows the propagation of proper collision response back onto the mechanical mesh.

@@

Because of the bending property of the shells relatively few elements are
necessary to simulate curved surface. To enrich the visual experience from
the simulated object it is desirable to use more triangles in the areas
with high curvature during rendering. 
Besides the visual accuracy the high resolution mesh correctly modeling the curved areas is needed for
interaction and correct simulation of collisions and constraints. 
To handle these, we also need to know how to map the velocities from the mechanical
mesh onto the high resolution mesh and how to map forces acting on the high
resolution mesh back onto the mechanical mesh.

@@

For contact modeling it is necessary to know the velocity of the contact points.
The values are directly available to us for the nodes of the mechanical mesh
but they can also be computed for any point on the surface.
Differentiating \eqref{eq-cubicbez2} by time we get:
%
\begin{equation}
    \dot{T} = \sum_{i=1}^{10} n_i \dot{P_i} = \sum_{i=1}^{10} n_i V_i
\end{equation}

Which means we use the same expression to interpolate the velocity only by
substituting the velocities at control points. 
For the corner \CD{\sout{nodes} $\rightarrow$ control points}  we
already have the velocities $V_i$. For the internal \CD{\sout{nodes} $\rightarrow$ control points}  we need to compute
them. We again make use of the kinematic link between nodes and control points.
%solve the problem of computing the velocity of a point attached to the moving rigid body.
For the edge points we applied to motion of a rigid body:
%
\begin{eqnarray}
    V_j = V_i + \omega_i \times P_{i-j}
\end{eqnarray}
%
where $i$ is the index of the corner node the edge node is attached to and
$\omega_i$ is the angular velocity at the node. For the central node
the formula is based on the derivate of equation \eqref{eq-central2}:
%
\begin{eqnarray}
    V_{10} = \frac{1}{3} \sum_{i=1}^3 \left( V_i + \omega_i \times P_{i-10} \right)
\end{eqnarray}

\subsection{Projection of surface forces}

For physical interaction with the object one also requires that the
forces applied on the object's surface are transmitted back as forces
and torques on the vertices of the mechanical mesh.
We compute the influence of force $F$ on the corners through all the
associated control points by the formula:
%
\begin{eqnarray}
    F_1 & = &  \sum_{i \in \{1,4,5\}} n_i F + \frac{n_{10}}{3} F \\
    F_2 & = &  \sum_{i \in \{2,6,7\}} n_i F + \frac{n_{10}}{3} F \\
    F_3 & = &  \sum_{i \in \{3,8,9\}} n_i F + \frac{n_{10}}{3} F 
\end{eqnarray}

Similarly we can compute the torques applied through edge control points and
the central control points:
%
\begin{gather}
    \begin{split}
        	\tau_1 = & P_{1-4} \times (n_4 F) + P_{1-5} \times (n_5 F) + \\
        & + \frac{1}{3} P_{1-10} \times (n_{10} F)
    \end{split}
    \\
    \begin{split}
       	\tau_2 = & P_{2-6} \times (n_6 F) + P_{2-7} \times (n_7 F) + \\
        & + \frac{1}{3} P_{2-10} \times (n_{10} F) \\
    \end{split}
    \\
    \begin{split}
        	\tau_3 = & P_{3-8} \times (n_8 F) + P_{3-9} \times (n_9 F) + \\
        & + \frac{1}{3} P_{3-10} \times (n_{10} F) \\
    \end{split}
\end{gather}

Note that these formula fulfill the physical principle of virtual power: $ P = \dot{T} \cdot F  = \sum_{i=1}^3 V_i \cdot F_i +  \omega_i  \cdot \tau_i$

@@

Because of the formulation of B\'ezier triangle is intrinsically present
there is a clean and obvious way of mapping high-resolution mesh for better
visualization. We also provide a technique of computing velocities of any point
on the surface and formula for back-projection of forces applied on the
surface onto the mechanical mesh. 
Those are essential components for the surgical simulators that need accurate contact geometry modeling without an overflow of computation need
The efficiency of the approach is demonstrated through convergence tests,  computation time benchmark and application to two different  simulation cases.
